# Harvest diameter

We need to match the data again for the diameter analysis because we must only consider trees for which the diameter is non-missing. So, we first select the trees for which diameter is non-missing, then we match the data using the same routine as in `matching.Rmd`. After matching, we will compare tree diameters. 

First, load the merged dataset. 

```{r}
library(tidyverse)
library(MatchIt)
library(weights)

load("data/df_merged.RData")
```

Then, match the data using the Genetic Matching algorithm. 

```{r}
# subset the data
df <- df_lagged %>%
  select(pltID, YEAR, spDIAremv_all, spDIAremv_ash, spDIAremv_nonash, EAB, PC1, pop_den_pm2, BAA_ash_county, BAA_nonash_county, 
         spDIAprev_ash, spDIAprev_nonash)

# ensure there are no missing values
df_nomiss <- df[complete.cases(df),]

# set seed for reproducibility
set.seed(9)

# matching
mod_match_dia <- MatchIt::matchit(EAB ~
                                BAA_ash_county + BAA_nonash_county + pop_den_pm2 + PC1
                                + spDIAprev_ash + spDIAprev_nonash,
          method = "genetic", data = df_nomiss, discard = "both", reestimate = T, pop.size=dim(df_nomiss)[1])

# assess the output
s.out_dia <- summary(mod_match_dia, standardize = T)
plot(s.out_dia, interactive=F)
summary(mod_match_dia, standardize = T)

# save a matched set
df_m_dia <- MatchIt::match.data(mod_match_dia)

save(df_m_dia, file = "data/df_matched_dia.RData")
```

Weighted t-test to infer statistical significance.

```{r}
# ash species
wtd.t.test(x=df_m_dia$spDIAremv_ash[df_m_dia$EAB==0],
           y=df_m_dia$spDIAremv_ash[df_m_dia$EAB==1],
           weight = df_m_dia$weights[df_m_dia$EAB==0],
           weighty= df_m_dia$weights[df_m_dia$EAB==1],
           bootse=T)

# non-ash species
wtd.t.test(x=df_m_dia$spDIAremv_nonash[df_m_dia$EAB==0],
           y=df_m_dia$spDIAremv_nonash[df_m_dia$EAB==1],
           weight = df_m_dia$weights[df_m_dia$EAB==0],
           weighty= df_m_dia$weights[df_m_dia$EAB==1],
           bootse=T)
```

Boxplots to visualize the differences in diameter.

```{r}
load("data/df_matched_dia.RData")

df_m_dia %>%
  select(EAB, spDIAremv_ash, spDIAremv_nonash) %>%
  mutate(EAB = ifelse(EAB==1, "EAB", "no EAB")) %>%
  rename(`Ash species` = spDIAremv_ash,
         `Non-ash species` = spDIAremv_nonash) %>%
  pivot_longer(c("Ash species", "Non-ash species"), names_to = "species", values_to = "diameter") %>%
  ggplot(aes(x=EAB, y=diameter, fill=EAB)) + geom_boxplot() +
  facet_wrap(~species) +
  ylab("harvest diameter (cm)") +
  xlab("") +
  theme_minimal() +
  scale_fill_manual(values = c("no EAB" = "#66c2a5",
                               "EAB" = "#fc8d62")) +
  theme(legend.title = element_blank(),
      legend.text = element_text(size=10),
      strip.text = element_text(size = 10),
      axis.title.y = element_text(size=10),
      axis.title.x = element_text(size=10),
      axis.text.x = element_text(size=10))
```
