# Model 1: Ash species

Load the matched data that we created in `matching.Rmd`.

```{r}
library(tidyverse)
library(MCMCvis)
library(rjags)
library(R2jags)
library(RColorBrewer)

load("data/df_matched.RData")
```

Create an ID variable.

```{r}
df_m$x <- as.numeric(rownames(df_m))
```

Replicate the data to incorporate weights.

```{r}
df_m$weights_round <- as.integer(round(df_m$weights, digits=1)*10)
df_m <- df_m[rep(row.names(df_m), df_m$weights_round),]
```

Prepare the model data.

```{r, echo=FALSE}
# subset the data
model_df <- subset(df_m, select = c(harvest_ash, spAGB_MgHa_prev_ash, percBAremv_ash, EAB, inv_period, pop_den, MHHI))

# ensure no missing values
model_df <- model_df[complete.cases(model_df),]

# scale variables
model_df <- model_df %>%
   mutate_at(c("pop_den", "MHHI"), funs(c(scale(.))))

# make the model matrix
X <- model.matrix(percBAremv_ash ~ spAGB_MgHa_prev_ash + EAB + pop_den + MHHI + inv_period, model_df)

# intercept
X[,1] <- rep(1, dim(X)[1])

# response variable
y <- model_df$percBAremv_ash

# n
n <- length(y)

# indicator variable (binary)
z <- model_df$harvest_ash

# the ones trick
ones <- rep(1, n)

# Beta distribution cannot handle ones
y[y==1] <- 0.999
```

Write a JAGS model. This is a zero-inflated beta regression model. 

```{r}
cat( "model{

  # For the ones trick
  C <- 10000

  for(i in 1:n){

    # define the logistic regression model, where w is the annual probability of NOT harvesting
    w[i] <- (1-j[i])^X[i,6]
    logit(j[i]) <- zeta[i]
    zeta[i] <- beta0 + beta1*X[i,2] + beta2*X[i,3] + beta3*X[i,4] + beta4*X[i,5]

    # define the gamma regression model for the mean. 
    eta[i] <- gamma0 + gamma1*X[i,2] + gamma2*X[i,3] + gamma3*X[i,4] + gamma4*X[i,5] 

    # for readability, define the log-likelihood of the gamma here
    logit(s[i]) <- eta[i]
    logGamma[i] <- log(dbeta(y[i], r*s[i], r*(1-s[i])))

    logLik[i] <- (1 - z[i]) * log(w[i]) + z[i] * ( log(1-w[i]) + logGamma[i])

    Lik[i] <- exp(logLik[i])

    # Use the ones trick
    
    p[i] <- Lik[i] / C
   ones[i] ~ dbern(p[i]) 
  }

  # PRIORS
  beta0 ~ dnorm(0, 0.01)
  beta1 ~ dnorm(0, 0.01)
  beta2 ~ dnorm(0, 0.01)
  beta3 ~ dnorm(0, 0.01)
  beta4 ~ dnorm(0, 0.01)

  gamma0 ~ dnorm(0, 0.01)
  gamma1 ~ dnorm(0, 0.01)
  gamma2 ~ dnorm(0, 0.01)
  gamma3 ~ dnorm(0, 0.01)
  gamma4 ~ dnorm(0, 0.01)

  shape ~ dunif(0, 10)
  r ~ dgamma(0.1, 0.1)
  
}
    ", fill=T, file="zero_inflated.txt"
)

rjags::load.module("glm")
cutData   <- list(y = y, X = X, n = n, z = z, ones=ones)
parNames <- c('beta0','beta1', 'beta2', 'beta3', 'beta4', 'gamma0', 'gamma1', 'gamma2', 'gamma3', 'gamma4', 'shape','r')

grow_ash <- jags(data=cutData, param=parNames, n.chains = 3,
                n.iter=2000, n.burnin=200, model.file="zero_inflated.txt")
```

Next, generate predictions. 

Since we inflated sample size by 10, we will scale the standard deviation by sqrt(10). This would be problematic if the priors were informative (since informative priors can be interpreted as sample size), but in our case, the priors are non-informative. So, we can scale the standard deviation post-hoc and rest assured that we are approximating the true value. 

Probability of harvest:

```{r}
params <- MCMCsummary(grow_ash,
                      params = c('beta0','beta1', 'beta2', 'beta3', 'beta4','shape', 'r'),
                      round = 3)

params$sd_mod <- params$sd*sqrt(10)
params$upper_mod <- params$mean+1.96*params$sd_mod
params$lower_mod <- params$mean-1.96*params$sd_mod

predictions <- data.frame(intercept = 1,
                       fracAGB_ash = rep(seq(0, 30, length.out=100),2),
                       EAB = c(rep(0,100), rep(1,100)),
                       pop_den = 0,
                       MHHI = 0,
                       inv_period = 1)

predictions$intercept <- params$mean[1]
predictions$ABG_mean <- params$mean[2]
predictions$EAB_mean <- params$mean[3]
predictions$EAB_5CI <- params$lower_mod[3]
predictions$EAB_95CI <- params$upper_mod[3]
predictions$pop_den_mean <- params$mean[4]
predictions$MHHI_mean <- params$mean[5]
predictions$shape <- params$mean[6]


for (i in 1:dim(predictions)[1]) {
  predictions$zeta[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_mean[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i]
  
  predictions$y[i] <- 1/(1+exp(-predictions$zeta[i]))
  
    predictions$zeta_5CI[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_5CI[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i] 
  
  predictions$y_5CI[i] <- 1/(1+exp(-predictions$zeta_5CI[i]))
  
    predictions$zeta_95CI[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_95CI[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i]
  
  predictions$y_95CI[i] <- 1/(1+exp(-predictions$zeta_95CI[i]))
}

df_plot <- predictions %>%
  mutate(EAB = ifelse(EAB==1, "EAB detected", "EAB not detected"))

ggplot(df_plot, aes(x=fracAGB_ash, y=y, group=EAB)) + geom_line(aes(color=EAB)) + 
  geom_ribbon(aes(ymin=y_5CI, ymax=y_95CI), alpha=0.4) +
  ylab("Annual probability of harvest") +
  xlab("aboveground biomass (MgHa)") +
  theme_minimal() + 
  scale_color_brewer(palette = "Dark2")
```

Intensity of harvest:

```{r}
params <- MCMCsummary(grow_ash,
                      params = c('gamma0','gamma1', 'gamma2', 'gamma3', 'gamma4','shape', 'r'),
                      round = 3)

params$sd_mod <- params$sd*sqrt(10)
params$upper_mod <- params$mean+1.96*params$sd_mod
params$lower_mod <- params$mean-1.96*params$sd_mod

predictions <- data.frame(intercept = 1,
                       fracAGB_ash = rep(seq(0, 30, length.out=100),2),
                       EAB = c(rep(0,100), rep(1,100)),
                       pop_den = 0,
                       MHHI = 0,
                       inv_period = 1)

predictions$intercept <- params$mean[1]
predictions$ABG_mean <- params$mean[2]
predictions$EAB_mean <- params$mean[3]
predictions$EAB_5CI <- params$lower_mod[3]
predictions$EAB_95CI <- params$upper_mod[3]
predictions$pop_den_mean <- params$mean[4]
predictions$MHHI_mean <- params$mean[5]
predictions$shape <- params$mean[6]
predictions$r <- params$mean[7]


for (i in 1:dim(predictions)[1]) {
  predictions$zeta[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_mean[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i]
  
    predictions$s[i] <- 1/(1+exp(-predictions$zeta[i]))
  predictions$y[i] <- qbeta(0.5, shape1 = predictions$r[i]*predictions$s[i], 
                            shape2 = predictions$r[i]*(1-predictions$s[i]))*100
  
    predictions$zeta_5CI[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_5CI[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i] 
  
      predictions$s[i] <- 1/(1+exp(-predictions$zeta_5CI[i]))
  predictions$y_5CI[i] <- qbeta(0.5, shape1 = predictions$r[i]*predictions$s[i], 
                            shape2 = predictions$r[i]*(1-predictions$s[i]))*100
  
    predictions$zeta_95CI[i] <- predictions$intercept[i] +
    predictions$ABG_mean[i]*predictions$fracAGB_ash[i] +
    predictions$EAB_95CI[i]*predictions$EAB[i] +
    predictions$pop_den_mean[i]*predictions$pop_den[i] +
    predictions$MHHI_mean[i]*predictions$MHHI[i]
  
  predictions$s[i] <- 1/(1+exp(-predictions$zeta_95CI[i]))
  predictions$y_95CI[i] <- qbeta(0.5, shape1 = predictions$r[i]*predictions$s[i], 
                            shape2 = predictions$r[i]*(1-predictions$s[i]))*100
}

df_plot <- predictions %>%
  mutate(EAB = ifelse(EAB==1, "EAB detected", "EAB not detected"))

ggplot(df_plot, aes(x=fracAGB_ash, y=y, group=EAB)) + geom_line(aes(color=EAB)) + 
  geom_ribbon(aes(ymin=y_5CI, ymax=y_95CI), alpha=0.4) +
  ylab("BA removed (%) ") +
  xlab("aboveground biomass (MgHa)") +
  theme_minimal() + 
  scale_color_brewer(palette = "Dark2")
```